---
title: "MATH 216 Homework 4"
author: "James Burke"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
library(tidyverse)
library(USAboundaries)
library(maptools)
library(sp)
library(broom)
```



## Admistrative:

Please indicate

* Who you collaborated with:
* Roughly how much time you spent on this HW so far: 7 Hours
* The URL of the RPubs published URL [here](http://rpubs.com/jamesburke4/HW-4).
* What gave you the most trouble:
* Any comments you have:



## Question 1:

**Question**:  DHL, a German shipping company, has a ton of money to build a
shipping hub in the United States. Where should they build it?

**Deliverable**: A choropleth map of the lower 48 states (no Puerto Rico) where
the color corresponds to the distance of the given area to the shipping hub,
which is marked on the map with a large red dot.

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
### PICKING LOCATION FOR SHIPPING HUB ###

# Extract Population Data
county_pop_data <- read_csv(file="County_Data.csv") %>%
  rename(
    Pop_Total = SE_T001_001,    # Total Population
    Pop_Density = SE_T002_002   # Population Density
    ) %>% 
  select(Geo_NAME, Geo_COUNTY, Geo_FIPS, Pop_Total, Pop_Density)

# Get County Spatial Polygons Data Frame
county_shp <- us_counties()

# Generate Centroid Values
centroids <- data.frame(coordinates(county_shp))

# Bind Centroid Values to County Data
county_data <- bind_cols(county_shp@data, centroids) %>% 
  rename(
    centroid_long = X1,
    centroid_lat = X2
  ) %>% 
  mutate(
    # creating geographic FIPS codes
    Geo_FIPS = as.numeric(paste(statefp, countyfp, sep = ""))
  )

# Joining Population Data with County Data
hub_data <- left_join(county_data, county_pop_data, by = "Geo_FIPS") %>%
  filter(!statefp %in% c("02", "15", "72")) # exclude Alaska, Hawaii, & Puerto Rico

# Manually Inputing Fairfax County, VA Population Data
hub_data$Pop_Total[hub_data$Geo_FIPS == 51600] = 1081726

hub_data <- hub_data %>%
  mutate(
    weighted_long = centroid_long * Pop_Total,
    weighted_lat = centroid_lat * Pop_Total
  ) %>% 
  select(Geo_FIPS, Pop_Total, weighted_long, weighted_lat) %>%
  filter(!is.na(Pop_Total))

# Calculating Hub Location - using center of mass formula
hub_x_coord <- sum(hub_data$weighted_long) / sum(hub_data$Pop_Total)
hub_y_coord <- sum(hub_data$weighted_lat) / sum(hub_data$Pop_Total)
```




```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
### CREATING CODE FOR MAP ###

# This code is repetitive but is generalized so that the map can easily be switched from
#  county data to state data and vice versa

# Generating State/County Dataset
chart_shp <- us_counties()    ## can substitute "us_states()" ##
centroids <- data.frame(coordinates(chart_shp))

chart_data <- bind_cols(chart_shp@data, centroids) %>% 
  rename(
    centroid_long = X1,
    centroid_lat = X2
  )

chart_polygon <- tidy(chart_shp, region="geoid")

geo_data <- left_join(chart_polygon, chart_data, by=c("id"="geoid")) %>%
  filter(!statefp %in% c("02", "15", "72"))

geo_data <- geo_data %>% 
  # Using distance formula to calculate distance from hub
  mutate(distance = sqrt(
                         (centroid_long - hub_x_coord)^2 + 
                         (centroid_lat - hub_y_coord)^2
                         ))
```



Map with shipping hub located at center of the US weighted by population. By using the center of mass formula and weighting the latitudes and longitudes of all US county centroids by the county populations according to the 2010 census. Fairfax County, VA population data was manually inputed.
```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
### PLOTTING THE MAP ###

ggplot(data=geo_data, aes(x=long, y=lat, group=group, fill=distance)) +
  # Plot choropleth polygons, low -> high being colored white -> dark green
  geom_polygon() +
  scale_fill_gradient(low="white", high="darkgreen") + 
  # Trace outlines of areas
  geom_path(size=0.1) +
  # Mark hub
  annotate("point", x=hub_x_coord, y=hub_y_coord, col="red", size=5) +
  # Correct aspect ratio for map
  coord_map()
```


Map with shipping hub location slightly adjusted for logistical purposes. The calculated hub was located in rural Missouri and just west of St. Louis. Therefore, I used Google Maps to find a location closer to the city of St. Louis where there is better access to major roads and airports, namely Lambert–St. Louis International Airport.
```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
hub_x_coord <- -90.384601
hub_y_coord <- 38.849079

ggplot(data=geo_data, aes(x=long, y=lat, group=group, fill=distance)) +
  geom_polygon() +
  scale_fill_gradient(low="white", high="darkgreen") +
  geom_path(size=0.1) +
  annotate("point", x=hub_x_coord, y=hub_y_coord, col="red", size=5) +
  coord_map()
```
